---
title: "wordle.qmd"
format: html
date: today
---

**## Step 1**
Create your own custom function called load_dictionary() that takes a single argument, “filename”, that can be used to read in either of the two data files your downloaded.
```{r}
library(tidyverse) # loading in required package for "read_tsv()"

# creating a function to read in downloaded data files
load_dictionary <- function(filename) { 
  read_tsv(filename, col_names = TRUE) # the argument "filename" will be used to easily load in the file path, as entering the URL in this function will pass it through the read_tsv() function. col_names is set to TRUE to ensure the first row of the file is the column names
}
```

Use the function to create two variables, solution_list and valid_list, that, respectively contain vectors of possible solution words and valid words to guess. That is, you should be able to run the following to create these two vectors:
```{r}
# creating vectors that contain the valid words to guess using load_dictionary() and passing in the file path
valid_list <- load_dictionary("/Users/sandanik/Documents/Repos/exercise-04/collins-scrabble-words-2019.txt")

# creating vectors that contain the possible solution words using load_dictionary() and passing in the file path
solution_list <- load_dictionary("/Users/sandanik/Documents/Repos/exercise-04/google-10000-english-usa-no-swears.txt")


# Checking to see if the function and vectors were made correctly
str(valid_list) # looks at structure of vector

str(solution_list) # looks at structure of vector
```

**## Step 2**
Winnow your variable solution_list to only include words that are included in valid_list. There are multiple ways that you could do this, but the set operation function, intersection() is an easy way. Use R help to look at the documentation for the intersection() function to see if you can get that to work. How many words are in your updated solution_list vector?